---
title: "Trends in incidence and outcomes of HIV-associated histoplasmosis in the Americas"
author: "Paridhi Ranadive"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    theme: cerulean
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: yes
---

<style type="text/css">
body, td{
  font-size: 14px;
}
code.r{
  font-size: 12px;
}
pre{
  font-size: 12px
}
div.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>


```{r setup, include=FALSE, warning =FALSE}
knitr::opts_chunk$set(echo = FALSE, message= FALSE, warning = FALSE, comment = NA)
options(knitr.kable.NA = '')
options(scipen = 999)
library(readr)
library(haven)
library(Hmisc)
library(dplyr)
library(tangram)
library(survival)
library(reshape2)
library(kableExtra)
library(survminer)
library(cmprsk)
library(rms)
library(psych)
library(gtsummary)
library(lattice)
library(survival)
library(purrr)
library(lubridate)
library(ggplot2)
library(sf)
library(countrycode)
library(rworldxtra)
library(maps)
library(RColorBrewer)
library(forestploter)
library(ggdist)
library(forcats)
library(stringr)


load(file="la.Rdata")
load(file="tn.Rdata")
load(file="incidence_histo.Rdata")
load(file="incidence_tn_histo.Rdata")
load(file="incid_b_histo.Rdata")
load(file="incid_m_histo.Rdata")
load(file="incid_b_long.Rdata")
load(file="incid_m_long.Rdata")
load(file="incid_long.Rdata")
load(file="incid_tn_long.Rdata")
load(file="incid_p_histo.Rdata")
load(file="incid_p_long.Rdata")
load(file="incid_h_histo.Rdata")
load(file="incid_h_long.Rdata")
load(file="incid_c_histo.Rdata")
load(file="incid_c_long.Rdata")
load(file="tab1.Rdata")
load(file="tab2.Rdata")
load(file="tab3.Rdata")
load(file="tab3_extra.Rdata")
load(file="tab4.Rdata")
load(file="tab5.Rdata")
load(file="tab6.Rdata")
load(file="tab7.Rdata")
load(file="tab8.Rdata")
load(file="tab9.Rdata")
load(file="tab10.Rdata")
load(file="tab11.Rdata")
load(file="tab12.Rdata")
load(file="tab13.Rdata")
load(file="tab14.Rdata")
load(file="cph_uni_tab_tn.Rdata")
load(file="cph_uni_tab_la.Rdata")
load(file="cph_summ.Rdata")
load(file="c.Rdata")
load(file="la_c.Rdata")
load(file="tn_c.Rdata")
load(file="cph_i.Rdata")


```

# Introduction
The aim of this project is to describe the incidence trends of histoplasmosis (excluding recurrent cases) in the Americas over time, compare trends in histoplasmosis incidence by occurrence before or after ART initiation, and examine risk1 factors for mortality following histoplasmosis diagnosis. It is hypothesized that incidence of histoplasmosis has decreased in the Americas, that risk1 for histoplasmosis will be substantially higher in the months immediately before and immediately after ART initiation, that mortality will be significantly higher among patients who develop histoplasmosis within 90 days of ART initiation and among patients in settings with endemic TB, and that trends in 30-day mortality after histoplasmosis diagnosis have decreased over time. Incidence of histoplasmosis will be examined per year from 2000-2021. We will examine trends using Poisson regression with splines to describe trends over time.


* Inclusion/ Exclusion Criteria: 

  + PWH who enrolled in care from January 1, 2000 to December 31, 2021, from cohorts with clinical endpoints available.  

  + Include all adults (age > 18 years)  

  + PWH with diagnoses of histoplasmosis prior to cohort entry will be excluded.  

  + Histoplasmosis diagnoses occurring on the same day as clinic entry(cohort entry) will be included (TBD: “same day” includes any diagnosis made the day of clinic entry - 30 days) 

  + Apply the VCCC inclusion criteria to CCASAnet data( 2 visits/lab results within a year of 1st visit) 
  + Excluded 56 ppl with male_y =8 in CCASAnet dataset and Intersexed in VCCC dataset

Pulmonary-only histoplasmosis cases are not routinely captured and will be excluded (TBD) 

* Notes:
   + Data set used for VCCC: histo_demographics_16jun2023.csv, histo_dx_17may2023.csv and histo_labs_21mar2023.csv
   + Data set used for CCASAnet: ccasanet_database_20230828
   + Included earliest diagnosis of Histoplasmosis
   + VCCC data has enrollment year as first provider visit and hence the baseline age is "age_at_first_visit"
   + ART Naive(Yes/No) defined from cohort entry(enrollment date)
   + ADE is defined as Yes for any ADE before 6 months and after 30 days of baseline
   + ADE for TN dataset are only for TB
   + Updated "ccasanet_database_20230828" dataset is used 
   + CD4 values greater 4000 were replaced with NAs
   + TB test is "Yes" if any of these test was : Culture positive tuberculosis, AFB Smear-positive,
   PCR positive,GeneXpert positive, MODS positive,Other diagnostic test results positive for CCASAnet
   + Tb treatment is "Yes" if tbtx_init is non missing
   

  
   
```{r flowchart, out.width="100%", out.height="100%"}                        
library(DiagrammeR)
DiagrammeR::grViz(

"digraph flowchart
                  {
      # node definitions with substituted label text
      node [fontname = Helvetica, shape = rectangle]        
      tab1 [label = '@@1']
      tab2 [label = '@@2']
      tab3 [label = '@@3']
      tab4 [label = '@@4']
      tab5 [label = '@@5']
      tab6 [label = '@@6']
      
      tab7 [label = '@@7']
      tab8 [label = '@@8']
      tab9 [label = '@@9']
      tab10 [label = '@@10']
      tab11 [label = '@@11']

      # edge definitions with the node IDs
      tab1 -> tab2 -> tab6 -> tab7 -> tab9 -> tab10
      tab1 -> tab3 -> tab4 -> tab5 -> tab8 -> tab11 
                  }
      
 [1]: 'All PWHIV enrolled at CCASAnet excluding Haiti and Argentina(n = 30844)\\n VCCC(n= 6870) '
      [2]: 'VCCC: Included(n = 6870)\\n Enrolled in care from January 1, 2000 to December 31, 2021, adults (age > 18 years)  '
      [3]: 'CCASAnet: Included(n = 26563) Exluded(n=4281)\\n Enrolled in care from January 1, 2000 to December 31, 2021, adults (age > 18 years) '
      [4]: 'Histoplasmisis cases for those enrolled in care from January 1, 2000 to December 31, 2021 and are adult  208\\n 50 Histo cases recorded prior to cohort entry'
      [5]: 'Included (n=26513) Histo cases(n=158) '
      [6]: 'Histoplasmosis cases 107(n=6870) '
      [7]: 'Included (n=6837): Histo cases(n=74)\\n 33 Histo cases recorded prior to cohort entry'
      [8]: 'VCCC Inclusion criteria on CCASAnet included (n=19886); Excluded (n=6627) \\n Histo cases (n=140) '
      [9]: 'Included(n=6837) with 74 Histo cases, 6763 without Histo'
      [10]: 'Included(n=6836);1 Excluded (birthsex= Intersexed) \\n 74 Histo cases, 6762 without Histo'
      [11]: 'Included(n=19836);50 Excluded (birthsex= Intersexed) \\n 140 Histo cases,19,696 without Histo'
      
                  
                  ")


```

# Descriptive Statistics
## CCASAnet {.tabset .tabset-pills}

### Table Overall 
```{r}
#la <- la %>% ungroup(patient_id)


ds <- la %>% select(baseline_y,age_b,risk1,sex,naive,follow,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,SITE,country_cat,mig_ende,birth_ende, ade_type,aids_first_y,l_alive_y,histo_y,death_y,death,rna_bb,rna_db,cd4_bb,cd4_db,histo,cp_y,tb,time_h,art_pre,art_post)
ds %>% 
  tbl_summary( 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```



### Table stratified by Site
```{r}
#la <- la %>% ungroup(patient_id)
ds <- la %>% select(baseline_y,age_b,risk1,sex,naive,follow,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,SITE,country_cat,mig_ende,birth_ende,ade_type,aids_first_y,l_alive_y,histo_y,death_y,death,rna_bb,rna_db,cd4_bb,cd4_db,histo,tb,cp_y,time_h,,art_pre,art_post)
ds %>% 
  tbl_summary( 
     by = SITE, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4","stat_5") ~ " Study Sites") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```


### Table stratified by Histoplasmosis 
```{r}
#la <- la %>% ungroup(patient_id)
ds <- la %>% select(baseline_y,age_b,risk1,sex,naive,follow,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,SITE,country_cat,mig_ende,birth_ende,ade_type,aids_first_y,l_alive_y,histo_y,death_y,death,rna_bb,rna_db,cd4_bb,cd4_db,histo,tb,cp_y,time_h,art_pre,art_post,pre_post_histo)
ds %>% 
  tbl_summary( 
     by = histo, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  add_p() %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ " Histoplasmosis ") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

### Table for Histo diagnisis stratified by ART status

```{r}
#la <- la %>% ungroup(patient_id)


ds <- la %>% select(baseline_y,age_b,risk1,sex,naive,follow,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,SITE,country_cat,mig_ende,birth_ende,ade_type,aids_first_y,l_alive_y,histo_y,death_y,death,rna_bb,rna_db,cd4_bb,cd4_db,histo,tb,cp_y,time_h,art_pre,art_post,pre_post_histo) %>% filter(histo=="Yes")
ds %>% 
  tbl_summary( 
     by = pre_post_histo, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  add_p() %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ " Histoplasmosis ") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

### Table for TB stratified by Histo


```{r}
#la <- la %>% ungroup(patient_id)


ds <- la %>% select(tb,cxpos_y,afbpos_y,pcrpos_y,genexpertpos_y,modspos_y,otherpos_y,tb_trt,tb_test,histo) %>% filter(tb=="Yes")
ds %>% 
  tbl_summary( 
     by = histo, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  add_p() %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ " Histoplasmosis ") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```
### Boxplots 
#### Boxplot for CD4(cell/mm^3)
```{r ,fig.align='center',fig.width=9,fig.height=7}
library(ggplot2)
library(gridExtra)
library(datasets)
library(grid)
g1 <-ggplot(la, aes(x= histo,y = cd4_b)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Baseline",
        x =" ", y = "CD4 cell/mm^3")

g2 <- ggplot(la, aes(x= histo,y = cd4_b_sq)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Square root transformed CD4 at baseline",
        x =" ", y = "CD4 cell/mm")



g3 <- ggplot(la, aes(x = factor(0),y = cd4_diag)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="At Histoplasmosis diagnosis",
        x =" ", y = "CD4 cell/mm^3")

g4 <- ggplot(la, aes(x = factor(0),y = cd4_diag_sq)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="Square root transformed CD4 at Histoplasmosis diagnosis",
        x =" ", y = "CD4 cell/mm")
grid.arrange(g1,g2,g3,g4, ncol = 2)
```

#### RNA(copies/ml)
```{r,fig.align='center',fig.width=9,fig.height=7}
library(ggplot2)
g1 <- ggplot(la, aes(x= histo,y = rna_b)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="HIV-1 RNA at baseline",
        x =" ", y = "copies/mL")

g2 <- ggplot(la, aes(x= histo,y = rna_b_log)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Log transformed HIV-1 RNA at baseline",
        x =" ", y = "copies/mL")

g3 <- ggplot(la, aes(x = factor(0),y = rna_diag)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="HIV-1 RNA at Histoplasmosis diagnosis",
        x =" ", y = "copies/mL")

g4 <- ggplot(la, aes(x = factor(0),y = rna_diag_log)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="Log transformed HIV-1 RNA at Histo diagnosis",
        x =" ", y = "copies/mL")
grid.arrange(g1,g2,g3,g4, ncol = 2)
```


## VCCC {.tabset .tabset-pills}
### Table
```{r}
tn <- tn %>%  ungroup(CFAR_PID)
ds_tn <- tn %>% select(year_of_enrollment,age_at_first_visit,age_at_histo,age_at_last_visit,age_at_death,death,risk1,sex,presentsex,country_cat,mig_ende,birth_ende,naive,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,AGE_AT_FIRST_INFECTION,AGE_AT_ART_START,rna_bb,rna_db,cd4_bb,cd4_db,histo,ade_type,time_h,follow_time_yrs,art_pre,art_post)
ds_tn %>% 
  tbl_summary( type = all_dichotomous() ~ "categorical") %>%
  bold_labels()

```


### Table stratified by Histoplasmosis 
```{r}
#tn <- tn %>% ungroup(CFAR_PID)
ds_tn <- tn %>% select(year_of_enrollment,age_at_first_visit,age_at_histo,age_at_last_visit,age_at_death,death,risk1,sex,presentsex,country_cat,mig_ende,birth_ende,naive,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,AGE_AT_FIRST_INFECTION,AGE_AT_ART_START,rna_bb,rna_db,cd4_bb,cd4_db,histo,ade_type,time_h,follow_time_yrs,art_pre,art_post)
ds_tn %>% 
  tbl_summary(
    by = histo, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ " Histoplasmosis ") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```



### Table for Histo cases stratified by pre-post ART 
```{r}
#tn <- tn %>% ungroup(CFAR_PID)
ds_tn <- tn %>% select(year_of_enrollment,age_at_first_visit,age_at_histo,age_at_last_visit,age_at_death,death,risk1,sex,presentsex,country_cat,mig_ende,birth_ende,naive,cd4_b,cd4_diag,rna_b,rna_b_log,rna_diag,AGE_AT_FIRST_INFECTION,AGE_AT_ART_START,rna_bb,rna_db,cd4_bb,cd4_db,histo,ade_type,time_h,follow_time_yrs,art_pre,art_post,pre_post_histo) %>% filter(histo== "Yes")
ds_tn %>% 
  tbl_summary(
    by = pre_post_histo, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ " Histoplasmosis ") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

### Boxlots
#### Boxplot for CD4(cell/mm^3)
```{r ,fig.align='center',fig.width=9,fig.height=7}
library(ggplot2)
library(gridExtra)
library(datasets)
library(grid)
g1 <-ggplot(tn, aes(x= histo,y = cd4_b)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Baseline",
        x =" ", y = "CD4 cell/mm^3")

g2 <- ggplot(tn, aes(x= histo,y = cd4_b_sq)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Square root transformed CD4 at baseline",
        x =" ", y = "CD4 cell/mm")



g3 <- ggplot(tn, aes(x = factor(0),y = cd4_diag)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="At Histoplasmosis diagnosis",
        x =" ", y = "CD4 cell/mm^3")

g4 <- ggplot(tn, aes(x = factor(0),y = cd4_diag_sq)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="Square root transformed CD4 at Histoplasmosis diagnosis",
        x =" ", y = "CD4 cell/mm")
grid.arrange(g1,g2,g3,g4, ncol = 2)
```

#### RNA at baseline
```{r,fig.align='center',fig.width=9,fig.height=7}
library(ggplot2)
g1 <- ggplot(tn, aes(x= histo,y = rna_b)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="HIV-1 RNA at baseline",
        x =" ", y = "copies/mL")

g2 <- ggplot(tn, aes(x= histo,y = rna_b_log)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
  theme_classic() +
  labs(title="Log transformed HIV-1 RNA at baseline",
        x =" ", y = "copies/mL")

g3 <- ggplot(tn, aes(x = factor(0),y = rna_diag)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="HIV-1 RNA at Histoplasmosis diagnosis",
        x =" ", y = "copies/mL")

g4 <- ggplot(tn, aes(x = factor(0),y = rna_diag_log)) +
    geom_boxplot() +
    stat_summary(fun.y = mean,
        geom = "point",
        size = 3,
        color = "steelblue") +
    theme_classic() +
  labs(title="Log transformed HIV-1 RNA at Histo diagnosis",
        x =" ", y = "copies/mL")
grid.arrange(g1,g2,g3,g4, ncol = 2)
```



# Incidents of histoplasmosis 

* Notes

  + Histoplasmosis diagnoses occurring on the same day as clinic entry will be included (TBD: “same day” includes any diagnosis made the day of clinic entry - 30 days) 
  + Only including cases with histo diagnosis after enrollment.
  + The point estimates are from the raw data with CI computed using year of diagnosis as a categorical variable for the modified Poisson model. Model summary for splines are also plotting along with point estimates.

## CCASAnet {.tabset .tabset-pills}
### Incidents
```{r}
kable(incidence_histo, format = "html", digits=2,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years"))  %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)


```
<br>

### Plots 

```{r,out.width="100%", out.height="100%"}
library(MASS)


incidence_histo$histo_y <- as.numeric(incidence_histo$histo_y)
incidence_histo_plot <- head(incidence_histo,-1)

## 

incidence_histo_plot$se_m <- with(incidence_histo_plot,sqrt((1-(n/py_follow_up))/n))
incidence_histo_plot$upl <- with(incidence_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
incidence_histo_plot$lcl <- with(incidence_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 4), offset= log(value),family = poisson(link = "log"), data = incid_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_long$year.cat <- as.factor(incid_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_long)
# incid_long$log_value <- log(incid_long$value)
# mod_ycat <- glm(histo ~ year.cat, offset = log(value) , family = negative.binomial(64), data=incid_long)
# mod_ycat <- glm.nb(histo ~ year.cat , data=incid_long)

mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:22) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 1:22) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$year <- c(2000:2021)

incidence_histo_plot <- cbind(incidence_histo_plot,prob1,ll,ul)
incidence_histo_plot <- merge(incidence_histo_plot,mod_ycat.mod , by.x="histo_y", by.y="year")

# plot(rep(incidence_histo_plot$histo_y,2),log10(c(incidence_histo_plot$ll,incidence_histo_plot$ul)),type = "n",axes = FALSE,xlab = "Year", ylab = "Incidence (per 1000)")
# axis(1)
# axis(2,at= c(-2,0,2),labels = c(0.01,1,100))
# box()
# lines(incidence_histo_plot$histo_y,log10(incidence_histo_plot$ll),lty =2)
# lines(incidence_histo_plot$histo_y,log10(incidence_histo_plot$ul),lty =2)
# lines(incidence_histo_plot$histo_y,log10(incidence_histo_plot$prob1))
# points(incidence_histo_plot$histo_y,log10(incidence_histo_plot$incidence_1000py))


mylabs <- c(0.001,0.01,0.1,1,10,100,1000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

ggplot(incidence_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray") +
  geom_line(aes(linetype="Model estimates")) +
  scale_x_continuous(breaks=c(2000, 2005, 2010, 2015, 2020)) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3)),labels=c(mylabs),limits = c(-2,2)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (CCASAnet)") +
  geom_line(aes(y=log10(incidence_1000py), linetype="Crude estimates")) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lower), ymax=log10(upper)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),
  plot.title= element_text(hjust=0.5), legend.position = "bottom")


```


### Point estimates and CI 
```{r}
test <- incidence_histo_plot %>% dplyr::select(incidence_1000py,upl,lcl,prob1,ll,ul,est.exp,lower,upper)

kable(test, format = "html", digits=2,
      col.names = c("Crude est", "Crude est ll", "Crude est ul", "Spline est.", "Spline ll","Spline ul", "Year as categ. est", "ll","ul")) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)

```

### Incidents and plots for Brazil 

+  Site specific plot's CI was computed using Wilson method for point estimates that had 0 incidence in a  calendar year. The value of 0 point estimate and lower CI were plotted at 0.001.


```{r brazil,out.width="100%", out.height="100%"}
library(MASS)
kable(incid_b_histo, format = "html", digits=2,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years","Point est", "lower","Upper"))  %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)



incid_b_histo$histo_y <- as.numeric(incid_b_histo$histo_y)
incid_b_histo_plot <- head(incid_b_histo,-1)

## 

incid_b_histo_plot$se_m <- with(incid_b_histo_plot,sqrt((1-(n/py_follow_up))/n))
incid_b_histo_plot$upl <- with(incid_b_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
incid_b_histo_plot$lcl <- with(incid_b_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 4), offset= log(value),family = poisson(link = "log"), data = incid_b_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_b_long$year.cat <- as.factor(incid_b_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_b_long)
# incid_long$log_value <- log(incid_long$value)
# mod_ycat <- glm(histo ~ year.cat, offset = log(value) , family = negative.binomial(64), data=incid_long)
# mod_ycat <- glm.nb(histo ~ year.cat , data=incid_long)

mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:22) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 1:22) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$year <- c(2000:2021)

incid_b_histo_plot <- cbind(incid_b_histo_plot,prob1,ll,ul)
incid_b_histo_plot <- merge(incid_b_histo_plot,mod_ycat.mod , by.x="histo_y", by.y="year")


mylabs <- c(0.001,0.01,0.1,1,10,100,1000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

ggplot(incid_b_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray")  +
  geom_line(aes(linetype="Model estimates")) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3)),labels=c(mylabs),limits = c(-3,3)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (Brazil)") +
  geom_point(aes(y=log10(incid_b_1000py))) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lcl), ymax=log10(upl)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),plot.title= element_text(hjust=0.5), legend.position = "bottom")
```

### Incidents and plots for Mexico

```{r mexico,out.width="100%", out.height="100%"}

library(MASS)

kable(incid_m_histo, format = "html", digits=2,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years","Point est", "lower","Upper"))  %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)



incid_m_histo$histo_y <- as.numeric(incid_m_histo$histo_y)
incid_m_histo_plot <- head(incid_m_histo,-1)

## 

incid_m_histo_plot$se_m <- with(incid_m_histo_plot,sqrt((1-(n/py_follow_up))/n))
incid_m_histo_plot$upl <- with(incid_m_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
incid_m_histo_plot$lcl <- with(incid_m_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 4), offset= log(value),family = poisson(link = "log"), data = incid_m_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_m_long$year.cat <- as.factor(incid_m_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_m_long)
# incid_long$log_value <- log(incid_long$value)
# mod_ycat <- glm(histo ~ year.cat, offset = log(value) , family = negative.binomial(64), data=incid_long)
# mod_ycat <- glm.nb(histo ~ year.cat , data=incid_long)

mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:22) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 2:22) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod <- mod_ycat.mod %>% filter(upper != Inf)
mod_ycat.mod$histo_y <- c(2002,2003,2004,2005,2006,2009,2010,2011,2013,2014,2016,2018,2019,2021)


histo_y <-  data.frame(histo_y = c(2000:2021),value = rep(1))
mexico_mod <- cbind(histo_y,prob1,ll,ul)
incid_m_histo_plot <- full_join(incid_m_histo_plot,mexico_mod,by ="histo_y")
incid_m_histo_plot <- full_join(incid_m_histo_plot,mod_ycat.mod,by ="histo_y")

incid_m_histo_plot <- incid_m_histo_plot %>% arrange(histo_y)
incid_m_histo_plot$incid_m_1000py <- as.numeric(incid_m_histo_plot$incid_m_1000py)
mylabs <- c(0.000,0.01,0.1,1,10,100,1000,10000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

incid_m_histo_plot$upl <- with(incid_m_histo_plot,ifelse(is.na(upl),Upper,upl))
incid_m_histo_plot$lcl <- with(incid_m_histo_plot,ifelse(is.na(lcl),Lower,lcl))
incid_m_histo_plot$lcl <- round(incid_m_histo_plot$lcl,digits=3)
incid_m_histo_plot$lcl <- ifelse(incid_m_histo_plot$lcl == 0.000,0.001,incid_m_histo_plot$lcl)
incid_m_histo_plot$lcl <- round(incid_m_histo_plot$lcl,digits=3) # again this step because it was not plotting
incid_m_histo_plot$incid_m_1000py <- round(incid_m_histo_plot$incid_m_1000py,digits=3) 
incid_m_histo_plot$incid_m_1000py <- ifelse(incid_m_histo_plot$incid_m_1000py == 0.000,0.001,incid_m_histo_plot$incid_m_1000py)


ggplot(incid_m_histo_plot,aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray") +
  geom_line(aes(linetype="Model estimates")) +
  scale_x_continuous(breaks=c(2002, 2005, 2010, 2015, 2020)) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3,4)),labels=c(mylabs),limits = c(-3,4)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (Mexico)") +
  geom_point(aes(y=log10(incid_m_1000py), linetype="Crude estimates")) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lcl), ymax=log10(upl)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14),plot.title= element_text(hjust=0.5), legend.position = "bottom")
```


### Indicents and plot  for Honduras

```{r honduras,out.width="100%", out.height="100%"}
library(MASS)

kable(incid_h_histo, format = "html", digits=2,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years","Point est", "lower","Upper"))  %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)


incid_h_histo$histo_y <- as.numeric(incid_h_histo$histo_y)
incid_h_histo_plot <- head(incid_h_histo,-1)

## 

incid_h_histo_plot$se_m <- with(incid_h_histo_plot,sqrt((1-(n/py_follow_up))/n))
incid_h_histo_plot$upl <- with(incid_h_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
incid_h_histo_plot$lcl <- with(incid_h_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 4), offset= log(value),family = poisson(link = "log"), data = incid_h_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_h_long$year.cat <- as.factor(incid_h_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_h_long)
# incid_long$log_value <- log(incid_long$value)
# mod_ycat <- glm(histo ~ year.cat, offset = log(value) , family = negative.binomial(64), data=incid_long)
# mod_ycat <- glm.nb(histo ~ year.cat , data=incid_long)

mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:22) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 2:22) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod <- mod_ycat.mod %>% filter(upper != Inf)
mod_ycat.mod$year <- c(2002,2003,2004,2005,2006,2007,2008,2012,2013,2015,2021)

histo_y <-  data.frame(histo_y = c(2000:2021),value = rep(1))
mod <- cbind(histo_y,prob1,ll,ul)
incid_h_histo_plot <- full_join(incid_h_histo_plot,mod,by ="histo_y")
#incid_h_histo_plot <- full_join(incid_h_histo_plot,mod_ycat.mod,by ="histo_y")

incid_h_histo_plot <- incid_h_histo_plot %>% arrange(histo_y)
incid_h_histo_plot$incid_h_1000py <- as.numeric(incid_h_histo_plot$incid_h_1000py)
incid_h_histo_plot$prob1 <- round(incid_h_histo_plot$prob1,digits=3)


mylabs <- c(0.000,0.01,0.1,1,10,100,1000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

incid_h_histo_plot$upl <- with(incid_h_histo_plot,ifelse(is.na(upl),Upper,upl))
incid_h_histo_plot$lcl <- with(incid_h_histo_plot,ifelse(is.na(lcl),Lower,lcl))
incid_h_histo_plot$lcl <- round(incid_h_histo_plot$lcl,digits=3)
incid_h_histo_plot$lcl <- ifelse(incid_h_histo_plot$lcl == 0.000,0.001,incid_h_histo_plot$lcl)
incid_h_histo_plot$lcl <- round(incid_h_histo_plot$lcl,digits=3) # again this step because it was not plotting
incid_h_histo_plot$incid_h_1000py <- round(incid_h_histo_plot$incid_h_1000py,digits=3) 
incid_h_histo_plot$incid_h_1000py <- ifelse(incid_h_histo_plot$incid_h_1000py == 0.000,0.001,incid_h_histo_plot$incid_h_1000py)

ggplot(incid_h_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray")  +
  geom_line(aes(linetype="Model estimates")) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3)),labels=c(mylabs),limits = c(-3,3)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (Honduras)") +
  geom_point(aes(y=log10(incid_h_1000py))) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lcl), ymax=log10(upl)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),plot.title= element_text(hjust=0.5), legend.position = "bottom")
```

### Incidents and plot for Peru

```{r peru,out.width="100%", out.height="100%"}
library(MASS)

kable(incid_p_histo, format = "html", digits=5,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years","Point est", "lower","Upper")) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)


incid_p_histo$histo_y <- as.numeric(incid_p_histo$histo_y)
incid_p_histo_plot <- head(incid_p_histo,-1)

## 

incid_p_histo_plot$se_m <- with(incid_p_histo_plot,sqrt((1-(n/py_follow_up))/n))
 incid_p_histo_plot$upl <- with(incid_p_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
 incid_p_histo_plot$lcl <- with(incid_p_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

CI_chile <- as.data.frame(binconf(incid_p_histo$n, incid_p_histo$num_persons, alpha=0.05,
       method="wilson",
      include.x=FALSE, include.n=FALSE, return.df=FALSE))

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 2), offset= log(value),family = poisson(link = "log"), data = incid_p_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2020),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2020),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_p_long$year.cat <- as.factor(incid_p_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_p_long)
mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:20) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 1:20) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod <- mod_ycat.mod %>% filter(upper != Inf)
mod_ycat.mod <- mod_ycat.mod[-1,]
mod_ycat.mod$histo_y <- c(2004,2005,2008,2009,2010,2011,2014,2015)

histo_y <-  data.frame(histo_y = c(2000:2020),value = rep(1))
mod <- cbind(histo_y,prob1,ll,ul)
incid_p_histo_plot <- full_join(incid_p_histo_plot,mod,by ="histo_y")
incid_p_histo_plot <- full_join(incid_p_histo_plot,mod_ycat.mod,by ="histo_y")

incid_p_histo_plot <- incid_p_histo_plot %>% arrange(histo_y)
incid_p_histo_plot$incid_p_1000py <- as.numeric(incid_p_histo_plot$incid_p_1000py)
incid_p_histo_plot$prob1 <- round(incid_p_histo_plot$prob1,digits=3)

mylabs <- c(0.000,0.01,0.1,1,10,100,1000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

incid_p_histo_plot$upl <- with(incid_p_histo_plot,ifelse(is.na(upl),Upper,upl))
incid_p_histo_plot$lcl <- with(incid_p_histo_plot,ifelse(is.na(lcl),Lower,lcl))
incid_p_histo_plot$lcl <- round(incid_p_histo_plot$lcl,digits=3)
incid_p_histo_plot$lcl <- ifelse(incid_p_histo_plot$lcl == 0.000,0.001,incid_p_histo_plot$lcl)
incid_p_histo_plot$lcl <- round(incid_p_histo_plot$lcl,digits=3) # again this step because it was not plotting
incid_p_histo_plot$incid_p_1000py <- round(incid_p_histo_plot$incid_p_1000py,digits=3) 
incid_p_histo_plot$incid_p_1000py <- ifelse(incid_p_histo_plot$incid_p_1000py == 0.000,0.001,incid_p_histo_plot$incid_p_1000py)
incid_p_histo_plot <- incid_p_histo_plot %>% filter(histo_y < 2020)


ggplot(incid_p_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray")  +
  geom_line(aes(linetype="Model estimates")) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3)),labels=c(mylabs),limits = c(-3,3)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (Peru)") +
  geom_point(aes(y=log10(incid_p_1000py))) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lcl), ymax=log10(upl)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),plot.title= element_text(hjust=0.5), legend.position = "bottom")

```


### Incidents and plots for Chile

```{r Chile,out.width="100%", out.height="100%"}
library(MASS)


kable(incid_c_histo, format = "html", digits=5,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years","Point est", "lower","Upper")) %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)


incid_c_histo$histo_y <- as.numeric(incid_c_histo$histo_y)
incid_c_histo_plot <- head(incid_c_histo,-1)

## 

incid_c_histo_plot$se_m <- with(incid_c_histo_plot,sqrt((1-(n/py_follow_up))/n))
 incid_c_histo_plot$upl <- with(incid_c_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
 incid_c_histo_plot$lcl <- with(incid_c_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

CI_chile <- as.data.frame(binconf(incid_c_histo$n, incid_c_histo$num_persons, alpha=0.05,
       method="wilson",
      include.x=FALSE, include.n=FALSE, return.df=FALSE))

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 2), offset= log(value),family = poisson(link = "log"), data = incid_c_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_c_long$year.cat <- as.factor(incid_c_long$year)

mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_c_long)
mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:21) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 1:21) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod <- mod_ycat.mod %>% filter(upper != Inf)
mod_ycat.mod$histo_y <- c(2012,2016,2019)

histo_y <-  data.frame(histo_y = c(2000:2021),value = rep(1))
chile_mod <- cbind(histo_y,prob1,ll,ul)
incid_c_histo_plot <- full_join(incid_c_histo_plot,chile_mod,by ="histo_y")
incid_c_histo_plot <- full_join(incid_c_histo_plot,mod_ycat.mod,by ="histo_y")

incid_c_histo_plot <- incid_c_histo_plot %>% arrange(histo_y)
incid_c_histo_plot$incid_c_1000py <- as.numeric(incid_c_histo_plot$incid_c_1000py)
incid_c_histo_plot$prob1 <- round(incid_c_histo_plot$prob1,digits=3)

mylabs <- c(0.00,0.01,0.1,1,10,100)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

incid_c_histo_plot$upl <- with(incid_c_histo_plot,ifelse(is.na(upl),Upper,upl))
incid_c_histo_plot$lcl <- with(incid_c_histo_plot,ifelse(is.na(lcl),Lower,lcl))
incid_c_histo_plot$lcl <- round(incid_c_histo_plot$lcl,digits=3)
incid_c_histo_plot$lcl <- ifelse(incid_c_histo_plot$lcl == 0.000,0.001,incid_c_histo_plot$lcl)
incid_c_histo_plot$lcl <- round(incid_c_histo_plot$lcl,digits=3) # again this step because it was not plotting
incid_c_histo_plot$incid_c_1000py <- round(incid_c_histo_plot$incid_c_1000py,digits=3) 
incid_c_histo_plot$incid_c_1000py <- ifelse(incid_c_histo_plot$incid_c_1000py == 0.000,0.001,incid_c_histo_plot$incid_c_1000py)

incid_c_histo_plot <- incid_c_histo_plot %>% filter(histo_y >= 2010)

ggplot(incid_c_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray")  +
  geom_line(aes(linetype="Model estimates")) +
  scale_x_continuous(breaks=c(2010, 2015, 2020)) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2)),labels=c(mylabs),limits = c(-3,2)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (Chile)") +
  geom_point(aes(y=log10(incid_c_1000py))) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lcl), ymax=log10(upl)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),plot.title= element_text(hjust=0.5), legend.position = "bottom")
```

## VCCC {.tabset .tabset-pills}

* Notes

  + Histoplasmosis diagnoses occurring on the same day as clinic entry will be included (TBD: “same day” includes any diagnosis made the day of clinic entry - 30 days)
  + First of July is taken as enrollment date, all other dates are calculated from .

### Incidents
```{r}
kable(incidence_tn_histo, format = "html", digits=2,
      col.names = c("Year", "Number of Histoplasmosis cases", "Number of persons", "Person-years of follow-up", "Incidence/1000 person-years"))  %>%
  kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)
```
<br>

### Plots 

```{r,out.width="100%", out.height="100%"}

incidence_tn_histo$histo_y <- as.numeric(incidence_tn_histo$histo_y)
incidence_tn_histo_plot <- head(incidence_tn_histo,-1)

incidence_tn_histo_plot$se_m <- with(incidence_tn_histo_plot,sqrt((1-(n/py_follow_up))/n))
incidence_tn_histo_plot$upl <- with(incidence_tn_histo_plot,exp(log(n/py_follow_up) + 1.96*se_m))*1000
incidence_tn_histo_plot$lcl <- with(incidence_tn_histo_plot,exp(log(n/py_follow_up) - 1.96*se_m))*1000

# Fit the quasi-Poisson model with natural splines
library(splines)
model <- glm(histo ~  ns(year, df = 4), offset = log(value),family = poisson(link = "log"), data = incid_tn_long)

# Summarize the model results
#summary(model)
nd <- data.frame(year = c(2000:2021),value = rep(1))
nd1 <- data.frame(year.cat = c(2000:2021),value = rep(1))
nd1$year.cat <- as.factor(nd1$year)
p <- predict(model, newdata = nd,se.fit = TRUE)
prob1 <- exp(p$fit)*1000
ll <- exp(p$fit - 1.96*p$se.fit)*1000
ul <- exp(p$fit + 1.96*p$se.fit)*1000
incid_tn_long$year.cat <- as.factor(incid_tn_long$year)
mod_ycat <- glm(histo ~ year.cat, offset= log(value), family=poisson(link = "log"), data=incid_tn_long)

mod_ycat.mod <- mod_ycat$coefficients
mod_ycat.mod <- as.data.frame(mod_ycat.mod)
names(mod_ycat.mod)[1] <- "est"
mod_ycat.mod$est.exp[1] <- exp(mod_ycat.mod$est[1])*1000 
for (i in 2:22) {
  mod_ycat.mod[, "est.exp"][i] <- exp(mod_ycat.mod[, "est"][1] + mod_ycat.mod[, "est"][i])*1000
}

#calculate standard errors
vcov.mod_ycat <- vcov(mod_ycat)
mod_ycat.mod$se <- NA
mod_ycat.mod$se[1] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[2, 2] + 2*(vcov(mod_ycat)[2, 1])
mod_ycat.mod$se[2] <- diag(vcov(mod_ycat))[1] + (vcov(mod_ycat))[3, 3] + 2*(vcov(mod_ycat)[3, 1])

for (i in 1:22) {
  mod_ycat.mod$se[i] <- sqrt(diag(vcov(mod_ycat))[1] + vcov(mod_ycat)[i, i] + 2*(vcov(mod_ycat)[1, i]))
  mod_ycat.mod$lower[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] -1.96*mod_ycat.mod$se[i])*1000
  mod_ycat.mod$upper[i] <- exp(mod_ycat.mod$est[1] + mod_ycat.mod$est[i] +1.96*mod_ycat.mod$se[i])*1000
}

mod_ycat.mod$se[1] <- sqrt(diag(vcov.mod_ycat))[1]
mod_ycat.mod$lower[1] <- exp(mod_ycat.mod$est[1] - 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$upper[1] <- exp(mod_ycat.mod$est[1] + 1.96*mod_ycat.mod$se[1])*1000
mod_ycat.mod$year <- c(2000:2021)

incidence_tn_histo_plot <- cbind(incidence_tn_histo_plot,prob1,ll,ul)
incidence_tn_histo_plot <- merge(incidence_tn_histo_plot,mod_ycat.mod , by.x="histo_y", by.y="year")

plot(rep(incidence_tn_histo_plot$histo_y,2),log10(c(incidence_tn_histo_plot$ll,incidence_tn_histo_plot$ul)),type = "n",axes = FALSE,xlab = "Year", ylab = "Incidence (per 1000)")
axis(1)
axis(2,at= c(-2,0,2),labels = c(0.01,1,100))
box()
lines(incidence_tn_histo_plot$histo_y,log10(incidence_tn_histo_plot$ll),lty =2)
lines(incidence_tn_histo_plot$histo_y,log10(incidence_tn_histo_plot$ul),lty =2)
lines(incidence_tn_histo_plot$histo_y,log10(incidence_tn_histo_plot$prob1))
points(incidence_tn_histo_plot$histo_y,log10(incidence_tn_histo_plot$incidence_tn_1000py))


mylabs <- c(0.001,0.01,0.1,1,10,100,1000)
#mylabs <- 10^(mylabs)
#mylabs <- round(mylabs, digits=1)

ggplot(incidence_tn_histo_plot, aes(histo_y, log10(prob1))) +
  geom_ribbon(aes(ymin= log10(ll), ymax= log10(ul)), fill="darkgray") +
  geom_line(aes(linetype="Model estimates")) +
  scale_x_continuous(breaks=c(2000, 2005, 2010, 2015, 2020)) +
  scale_y_continuous(breaks = c(c(-3,-2,-1,0,1,2,3)),labels=c(mylabs),limits = c(-2,2)) +
  labs(x="Year", y="Incidence estimates / 1000 person-years", title= "Incidence of Histoplasmosis diagnosis (VCCC)") +
  geom_line(aes(y=log10(incidence_tn_1000py), linetype="Crude estimates")) +
  scale_linetype_manual(name="", values=c("Model estimates"=1, "Crude estimates"=2)) +
  geom_errorbar(aes(ymin=log10(lower), ymax=log10(upper)), width=0.2, color="gray50") +
  theme_bw() +
  theme(
  axis.title.x = element_text(size = 16, color="black"),
  axis.text.x = element_text(size = 16, color="black"),
  axis.text.y = element_text(size = 16, color="black"),
  axis.title.y = element_text(size = 14, color="black"),plot.title= element_text(hjust=0.5), legend.position = "bottom")

```
 
<br>

# Map for count of people migrated from Histo endemic countries {.tabset .tabset-pills}

## CCASAnet 

```{r maps, warning=FALSE, echo=FALSE, message=FALSE}
#table_la <- la %>% filter(country_cat == "Yes")
table_la <- la %>% filter(mig_ende == "Yes")
table_la <- as.data.frame(table(table_la$origin))
table_la$birth_c <- as.character(table_la$Var1)

table_la$birth_c <- with(table_la, ifelse(birth_c == "OUTSIDE", NA, birth_c))

table_la$country <- countrycode(table_la$birth_c, origin = "iso3c", destination = "country.name")                            
table_la$country <- with(table_la, ifelse(country == "United States", "USA", country))
world_map <- map_data("world")
table_la$count_group <- cut(table_la$Freq, 
                            breaks = c(-Inf, 10, 50, 100,200, Inf), 
                            labels = c("Less than 10", "10-50", "50-100","100-200","More than 200"))

ggplot(table_la) +
  geom_map(data =table_la ,aes(map_id = country, fill = count_group), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'grey20', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Migrated from Histo Endemic  (CCASAnet)", fill ="Count") +
  #scale_fill_manual(name = "Counts", values = brewer.pal(6, name = "Dark2")) +
  theme_void() +
  coord_fixed()

```

## VCCC 

```{r maps2, warning=FALSE, echo=FALSE, message=FALSE}
#table_tn <- tn %>% filter(country_cat == "Yes")
table_tn <- tn %>% filter(mig_ende == "Yes")
table_tn <- as.data.frame(table(table_tn$birthcountry))
table_tn$birth_c <- as.character(table_tn$Var1)
table_tn$birth_c <- countrycode(table_tn$birth_c, origin = 'country.name', destination = 'iso3c')
table_tn$country <- countrycode(table_tn$birth_c, origin = "iso3c", destination = "country.name") 
table_tn$country <- with(table_tn, ifelse(country == "United States", "USA", country))
table_tn$count_group <- cut(table_tn$Freq, 
                            breaks = c(-Inf, 10, 50, Inf), 
                            labels = c("Less than 10", "10-50", "More than 50"))
world_map <- map_data("world")
ggplot(table_tn) +
  geom_map(data =table_tn ,aes(map_id = country, fill = count_group), map = world_map) +
  geom_polygon(data = world_map, aes(x = long, y = lat, group = group), colour = 'grey20', fill = NA) +
  expand_limits(x = world_map$long, y = world_map$lat) +
  labs(title = "Migrated from Histo Endemic  (VCCC)",fill ="Count") +
  #scale_fill_manual(name = "Counts", values = brewer.pal(5, name = "Dark2")) +
  theme_void() +
  coord_fixed()

```

<br>

# Regression Models (Poisson)
## Unadjusted and Adjusted model with follow-up time as offset for CCASAnet data  

 * Notes 

   + Follow-up time is defined as time from enrollment to first of l_alive_date or cohort end date 
   + Following model use age,cd4 and RNA as spline. 

```{r regression adjusted, warning=FALSE, echo=FALSE, message=FALSE} 
 colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE) 
 colnames(tab1) <- colnames 


 options(digits = 3) 
 options(scipen=999) 
 kbl(tab1, booktabs = TRUE) %>%  
   add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))  %>% 
   pack_rows(group_label = "Age (ref= 35)", start_row = 1, end_row = 1,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "CD4 (ref = 1000)", start_row = 9, end_row = 9,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "Gender (ref = Female)", start_row = 14, end_row = 14,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "HIV acquisition factor (ref = MSM)", start_row = 15, end_row = 15,label_row_css = "background-color: #666; color: #fff;") %>%  
    pack_rows(group_label = "rna (ref =Undetectable)", start_row = 18, end_row = 18,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "Site (ref =Brazil)", start_row = 19, end_row = 19,label_row_css = "background-color: #666; color: #fff;") 

```  



## Unadjusted and Adjusted model with follow-up time as offset for VCCC data  

```{r regression adjusted vccc new, warning=FALSE, echo=FALSE, message=FALSE} 
 colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE) 
 colnames(tab2) <- colnames 


 options(digits = 3) 
 options(scipen=999) 
 kbl(tab2, booktabs = TRUE) %>%  
   add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

```  


# Regression Models: time updated (Poisson)

* Notes
  + Time updated ART-status
  + Follow-up time is defined as time from enrollment to first of l_alive_date or cohort end date
  + Following models use age, cd4 and RNA as spline.
  + Those with missing ART start date are included in pre-ART initiation
  + VCCC doesn't have time updated ART in the drive
  + CCASAnet has 59 histo cases in pre-ART period and 87 cases in post-ART period
  + VCCC has 8 pre-ART period and 66 post-ART period.
  

## Unadjusted and Adjusted model with follow-up time as offset for CCASAnet data time updated ART
```{r regression n adjusted ccasanet, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab3) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab3, booktabs = TRUE) %>% 
  add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows(group_label = "Age (ref= 35)", start_row = 1, end_row = 1,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "CD4 (ref = 200)", start_row = 9, end_row = 9,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "log HIV RNA (ref= 5)", start_row = 13, end_row = 13,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Year (ref = 2010)", start_row = 16, end_row = 16,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Gender (ref = Female)", start_row = 20, end_row = 20,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "HIV acquisition factor (ref = MSM)", start_row = 21, end_row = 21,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "Site (ref =Brazil)", start_row = 24, end_row = 24,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "ART status (ref =post-ART)", start_row = 28, end_row = 28,label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows(group_label = "Migrated from Histo endemic country (ref =No)", start_row = 29, end_row = 29,label_row_css = "background-color: #666; color: #fff;")
  
``` 



## Unadjusted and Adjusted model with follow-up time as offset for VCCC data time updated ART
```{r regression adjusted vccc, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab4) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab4, booktabs = TRUE) %>% 
  add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
  
  
```



## Imputed model CCASAnet (baseline year)
```{r impu regression adjusted ccasanet, warning=FALSE, echo=FALSE, message=FALSE,fig.asp = 1,out.width="80%", out.height="80%"}

load(file="tab5.Rdata")
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab5) <- colnames



options(digits = 2)
options(scipen=999)
kbl(tab5, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

tab5$RR1 <- log(tab5$RR)
tab5$Lower_ci <- log(tab5$`Lower .95 CI`)
tab5$Upper_ci <- log(tab5$`Upper .95 CI`)

tab5$RR1 <- round(tab5$RR1,digits=2)
tab5$Lower_ci <- round(tab5$Lower_ci,digits=2)
tab5$Upper_ci <- round(tab5$Upper_ci,digits=2)

tab5 <- tab5[-c(2,4,5,7,9,24,30,32),]

tab5$var <- c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25 ","100","350","500",
                    " log HIV RNA (ref=5)","1", "3","7","Year (ref=2010)", "2000","2005","2015","2020","Sex: Male","HIV acquisition factor (ref =MSM)","Heterosexual ","IDU ", "Other/Unknown ", 
                    "ART initiation status: pre initiation", "Migration status: Yes","Country (ref=Brazil)", "Chile ", "Honduras ","Mexico ", "Peru ")


tab5$var <- factor(tab5$var, ordered = TRUE,levels = c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25 ","100","350","500",
                    " log HIV RNA (ref=5)","1", "3","7","Year (ref=2010)", "2000","2005","2015","2020","Sex: Male","HIV acquisition factor (ref =MSM)","Heterosexual ","IDU ", "Other/Unknown ", 
                    "ART initiation status: pre initiation", "Migration status: Yes","Country (ref=Brazil)", "Chile ", "Honduras ","Mexico ", "Peru "))

tab5 <- tab5[c(1,2,3,4,19,20,21,22,23,14,15,16,17,18,24,5:13,25:30),]

tab5 <- tab5 %>%
  mutate(label = ifelse(is.na(RR), "", sprintf("%.2f [%.2f, %.2f]", RR, `Lower .95 CI`, `Upper .95 CI`)))


mylabs <- c(0.0156,0.0625,0.25,1,4)

p1 <- ggplot(tab5, aes(x=factor (var, level=var), y=tab5$RR1, ymin=tab5$Lower_ci, ymax=tab5$Upper_ci)) + 
  geom_linerange() +
  geom_hline(aes(x=0,yintercept=log(1)), lty=2) +
  geom_point(shape=18) +
   
  scale_y_continuous(name= " Relative risk ratio 95% CI (log scale)",breaks = c(c(-6,-4,-2,0,2)),labels=c(mylabs)) +
  coord_flip() +
  theme(
    text = element_text(family = "Georgia"),
    plot.title = element_text(size = 14, color = "black", face = "bold"),
    plot.subtitle = element_text(face = "italic", color = "black", size = 14),
    plot.caption = element_text(face = "italic", size = 14, color = "black"),
    axis.title.x = element_text(face = "bold", size = 14, color = "black"),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "#fcfbfd"),
    plot.margin = unit(c(1, 1, 1, 0), "cm")) +
    scale_x_discrete(limits = rev, breaks = tab5$var,expand = expansion(add = 0)) 
    

# Text column
p2 <- ggplot(tab5, aes(y=factor (var, level=var),x = 1, label = label)) +
  geom_text(hjust = 0,size=3.6) +
  scale_y_discrete(limits = rev, breaks = var) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = unit(c(0.6, 1, 2.3, -4), "cm"))
# Combine plots
forest1 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 1))

 
``` 


## Imputed model CCASAnet (calender year)
```{r impu regression adjusted ccasanet new, warning=FALSE, echo=FALSE, message=FALSE,fig.asp = 1,out.width="80%", out.height="80%"}

load(file="tab5_extra.Rdata")
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab5_extra) <- colnames



options(digits = 2)
options(scipen=999)
kbl(tab5_extra, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

tab5_extra$RR1 <- log(tab5_extra$RR)
tab5_extra$Lower_ci <- log(tab5_extra$`Lower .95 CI`)
tab5_extra$Upper_ci <- log(tab5_extra$`Upper .95 CI`)

tab5_extra$RR1 <- round(tab5_extra$RR1,digits=2)
tab5_extra$Lower_ci <- round(tab5_extra$Lower_ci,digits=2)
tab5_extra$Upper_ci <- round(tab5_extra$Upper_ci,digits=2)

tab5_extra <- tab5_extra[-c(2,4,5,7,9,24,30,32),]

tab5_extra$var <- c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25 ","100","350","500",
                    " log HIV RNA (ref=5)","1", "3","7","Year (ref=2010)", "2000","2005","2015","2020","Sex: Male","HIV acquisition factor (ref =MSM)","Heterosexual ","IDU ", "Other/Unknown ", 
                    "ART initiation status: pre initiation", "Migration status: Yes","Country (ref=Brazil)", "Chile ", "Honduras ","Mexico ", "Peru ")


tab5_extra$var <- factor(tab5_extra$var, ordered = TRUE,levels = c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25 ","100","350","500",
                    " log HIV RNA (ref=5)","1", "3","7","Year (ref=2010)", "2000","2005","2015","2020","Sex: Male","HIV acquisition factor (ref =MSM)","Heterosexual ","IDU ", "Other/Unknown ", 
                    "ART initiation status: pre initiation", "Migration status: Yes","Country (ref=Brazil)", "Chile ", "Honduras ","Mexico ", "Peru "))

tab5_extra <- tab5_extra[c(1,2,3,4,19,20,21,22,23,14,15,16,17,18,24,5:13,25:30),]

tab5_extra <- tab5_extra %>%
  mutate(label = ifelse(is.na(RR), "", sprintf("%.2f [%.2f, %.2f]", RR, `Lower .95 CI`, `Upper .95 CI`)))


mylabs <- c(0.0156,0.0625,0.25,1,4)

p1 <- ggplot(tab5_extra, aes(x=factor (var, level=var), y=tab5_extra$RR1, ymin=tab5_extra$Lower_ci, ymax=tab5_extra$Upper_ci)) + 
  geom_linerange() +
  geom_hline(aes(x=0,yintercept=log(1)), lty=2) +
  geom_point(shape=18) +
   
  scale_y_continuous(name= " Relative risk ratio 95% CI (log scale)",breaks = c(c(-6,-4,-2,0,2)),labels=c(mylabs)) +
  coord_flip() +
  theme(
    text = element_text(family = "Georgia"),
    plot.title = element_text(size = 14, color = "black", face = "bold"),
    plot.subtitle = element_text(face = "italic", color = "black", size = 14),
    plot.caption = element_text(face = "italic", size = 14, color = "black"),
    axis.title.x = element_text(face = "bold", size = 14, color = "black"),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "#fcfbfd"),
    plot.margin = unit(c(1, 1, 1, 0), "cm")) +
    scale_x_discrete(limits = rev, breaks = tab5_extra$var,expand = expansion(add = 0)) 
    

# Text column
p2 <- ggplot(tab5_extra, aes(y=factor (var, level=var),x = 1, label = label)) +
  geom_text(hjust = 0,size=3.6) +
  scale_y_discrete(limits = rev, breaks = var) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = unit(c(0.6, 1, 2.3, -4), "cm"))
# Combine plots
forest1 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 1))

 
``` 

<br>

## Imputed model VCCC (baseline year)
```{r impu 2 regression adjusted ccasanet, warning=FALSE, echo=FALSE, message=FALSE,fig.asp = 1, out.width="80%", out.height="70%"}
load(file="tab6.Rdata")
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab6) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab6, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

tab6$RR1 <- log(tab6$RR)
tab6$Lower_ci <- log(tab6$`Lower .95 CI`)
tab6$Upper_ci <- log(tab6$`Upper .95 CI`)

tab6$RR1 <- round(tab6$RR1,digits=2)
tab6$Lower_ci <- round(tab6$Lower_ci,digits=2)
tab6$Upper_ci <- round(tab6$Upper_ci,digits=2)

tab6 <- tab6[-c(19),]

# tab6$RR <- with(tab6,ifelse(RR < -6 ,-4,RR ))
# tab6$`Lower .95 CI` <- with(tab6,ifelse(`Lower .95 CI` < -6 ,-4,`Lower .95 CI` ))
# tab6$`Upper .95 CI` <- with(tab6,ifelse(`Upper .95 CI` < -6 ,-4,`Upper .95 CI` ))

tab6$var <- c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25  ","  100 ","350","500  ","log HIV RNA(ref=  5 )", "1", "3" ,"7",
                    "Year (ref=2010)", "2000","2005","2015","2020","Sex: Male", "Risk factor (ref =MSM)","Hetero ","IDU ", "Other/Unknown ",
                    "ART initiation status (ref=post)","ART initiation status: pre initiation", "Migration status (ref=No)", "Migration status: Yes")

tab6$var <- factor(tab6$var, ordered = TRUE,levels = c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25  ","  100 ","350","500  ","log HIV RNA(ref=  5 )", "1", "3" ,"7",
                    "Year (ref=2010)", "2000","2005","2015","2020","Sex: Male", "Risk factor (ref =MSM)","Hetero ","IDU ", "Other/Unknown ",
                    "ART initiation status (ref=post)","ART initiation status: pre initiation", "Migration status (ref=No)","Migration status: Yes"))

tab6 <- tab6[c(1:4,19,20,21,22,23,14,15,16,17,18,25,5:13,27),]

tab6 <- tab6 %>%
  mutate(label = ifelse(is.na(RR), "", sprintf("%.2f [%.2f, %.2f]", RR, `Lower .95 CI`, `Upper .95 CI`)))

mylabs <- c(0.0156, 0.0625,0.25,1,4)
p1 <- ggplot(tab6, aes(x=factor (tab6$var, level=tab6$var), y=tab6$RR1, ymin=tab6$Lower_ci, ymax=tab6$Upper_ci)) + 
  geom_linerange() +
  geom_hline(aes(x=0, yintercept=log(1)), lty=2) +
  geom_point(shape=18) +
  scale_y_continuous(name= " Relative risk ratio 95% CI (log scale)",breaks = c(c(-6,-4,-2,0,2)),labels=c(mylabs)) +
  coord_flip() +
  theme(
    text = element_text(family = "Georgia"),
    plot.title = element_text(size = 14, color = "black", face = "bold"),
    plot.subtitle = element_text(face = "italic", color = "black", size = 14),
    plot.caption = element_text(face = "italic", size = 12, color = "black"),
    axis.title.x = element_text(face = "bold", size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "#fcfbfd"),
    plot.margin = unit(c(1, 1, 1, 0), "cm"))  +
  scale_x_discrete(limits = rev)
  
# Text column
p2 <- ggplot(tab6, aes(y=factor (tab6$var, level=tab6$var), x = 1, label = label)) +
  geom_text(hjust = 0,size=3.6) +
  scale_y_discrete(limits = rev, breaks = tab6$var) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = unit(c(1, 1, 2.3, -4), "cm"))
# Combine plots
forest2 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 1))


``` 


## Imputed model VCCC (calender year)
```{r impu VCCC extra regression adjusted ccasanet, warning=FALSE, echo=FALSE, message=FALSE,fig.asp = 1, out.width="80%", out.height="70%"}
load(file="tab6_extra.Rdata")
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab6_extra) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab6_extra, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

tab6_extra$RR1 <- log(tab6_extra$RR)
tab6_extra$Lower_ci <- log(tab6_extra$`Lower .95 CI`)
tab6_extra$Upper_ci <- log(tab6_extra$`Upper .95 CI`)

tab6_extra$RR1 <- round(tab6_extra$RR1,digits=2)
tab6_extra$Lower_ci <- round(tab6_extra$Lower_ci,digits=2)
tab6_extra$Upper_ci <- round(tab6_extra$Upper_ci,digits=2)

tab6_extra <- tab6_extra[-c(19),]

# tab6_extra$RR <- with(tab6_extra,ifelse(RR < -6 ,-4,RR ))
# tab6_extra$`Lower .95 CI` <- with(tab6_extra,ifelse(`Lower .95 CI` < -6 ,-4,`Lower .95 CI` ))
# tab6_extra$`Upper .95 CI` <- with(tab6_extra,ifelse(`Upper .95 CI` < -6 ,-4,`Upper .95 CI` ))

tab6_extra$var <- c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25  ","  100 ","350","500  ","log HIV RNA(ref=  5 )", "1", "3" ,"7",
                    "Year (ref=2010)", "2000","2005","2015","2020","Sex: Male", "Risk factor (ref =MSM)","Hetero ","IDU ", "Other/Unknown ",
                    "ART initiation status (ref=post)","ART initiation status: pre initiation", "Migration status (ref=No)", "Migration status: Yes")

tab6_extra$var <- factor(tab6_extra$var, ordered = TRUE,levels = c("Age (ref=35)","25","45","55",
                    "CD4 (ref=200)","25  ","  100 ","350","500  ","log HIV RNA(ref=  5 )", "1", "3" ,"7",
                    "Year (ref=2010)", "2000","2005","2015","2020","Sex: Male", "Risk factor (ref =MSM)","Hetero ","IDU ", "Other/Unknown ",
                    "ART initiation status (ref=post)","ART initiation status: pre initiation", "Migration status (ref=No)","Migration status: Yes"))

tab6_extra <- tab6_extra[c(1:4,19,20,21,22,23,14,15,16,17,18,25,5:13,27),]

tab6_extra <- tab6_extra %>%
  mutate(label = ifelse(is.na(RR), "", sprintf("%.2f [%.2f, %.2f]", RR, `Lower .95 CI`, `Upper .95 CI`)))

mylabs <- c(0.0156, 0.0625,0.25,1,4)
p1 <- ggplot(tab6_extra, aes(x=factor (tab6_extra$var, level=tab6_extra$var), y=tab6_extra$RR1, ymin=tab6_extra$Lower_ci, ymax=tab6_extra$Upper_ci)) + 
  geom_linerange() +
  geom_hline(aes(x=0, yintercept=log(1)), lty=2) +
  geom_point(shape=18) +
  scale_y_continuous(name= " Relative risk ratio 95% CI (log scale)",breaks = c(c(-6,-4,-2,0,2)),labels=c(mylabs)) +
  coord_flip() +
  theme(
    text = element_text(family = "Georgia"),
    plot.title = element_text(size = 14, color = "black", face = "bold"),
    plot.subtitle = element_text(face = "italic", color = "black", size = 14),
    plot.caption = element_text(face = "italic", size = 12, color = "black"),
    axis.title.x = element_text(face = "bold", size = 12, color = "black"),
    axis.title.y = element_blank(),
    axis.text.x = element_text(size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "#fcfbfd"),
    plot.margin = unit(c(1, 1, 1, 0), "cm"))  +
  scale_x_discrete(limits = rev)
  
# Text column
p2 <- ggplot(tab6_extra, aes(y=factor (tab6_extra$var, level=tab6_extra$var), x = 1, label = label)) +
  geom_text(hjust = 0,size=3.6) +
  scale_y_discrete(limits = rev, breaks = tab6_extra$var) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = unit(c(1, 1, 2.3, -4), "cm"))
# Combine plots
forest2 <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 1))


``` 


<br>

# Regression model Pre ART 

## Unadjusted and Adjusted model with follow-up time as offset for CCASAnet data time updated ART (only pre ART)
```{r regression n adjusted ccasanet pre, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab7) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab7, booktabs = TRUE) %>% 
  add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows(group_label = "Age (ref= 35)", start_row = 1, end_row = 1,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "CD4 (ref = 200)", start_row = 9, end_row = 9,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "log HIV RNA (ref= 5)", start_row = 13, end_row = 13,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Year (ref = 2010)", start_row = 16, end_row = 16,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Gender (ref = Female)", start_row = 20, end_row = 20,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "HIV acquisition factor (ref = MSM)", start_row = 21, end_row = 21,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "Site (ref =Brazil)", start_row = 24, end_row = 24,label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows(group_label = "Migrated from Histo endemic country (ref =No)", start_row = 28, end_row = 28,label_row_css = "background-color: #666; color: #fff;")
  
``` 


## Imputed model CCASAnet (Pre ART)

```{r r impu regression adjusted ccasanet pre, warning=FALSE, echo=FALSE, message=FALSE}
 colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
 colnames(tab9) <- colnames


 options(digits = 3)
 options(scipen=999)
 kbl(tab9, booktabs = TRUE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))

```


# Regression models for post ART


## Unadjusted and Adjusted model with follow-up time as offset for CCASAnet data time updated ART (only post ART)
```{r regression n adjusted ccasanet post, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab11) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab11, booktabs = TRUE) %>% 
  add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  pack_rows(group_label = "Age (ref= 35)", start_row = 1, end_row = 1,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "CD4 (ref = 200)", start_row = 9, end_row = 9,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "log HIV RNA (ref= 5)", start_row = 13, end_row = 13,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Year (ref = 2010)", start_row = 16, end_row = 16,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "Gender (ref = Female)", start_row = 20, end_row = 20,label_row_css = "background-color: #666; color: #fff;") %>%
  pack_rows(group_label = "HIV acquisition factor (ref = MSM)", start_row = 21, end_row = 21,label_row_css = "background-color: #666; color: #fff;") %>% 
   pack_rows(group_label = "Site (ref =Brazil)", start_row = 24, end_row = 24,label_row_css = "background-color: #666; color: #fff;") %>%
pack_rows(group_label = "Migrated from Histo endemic country (ref =No)", start_row = 28, end_row = 28,label_row_css = "background-color: #666; color: #fff;")
  
``` 



## Unadjusted and Adjusted model with follow-up time as offset for VCCC data time updated ART (Post ART TN)
```{r regression adjusted vccc post, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value","RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab12) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab12, booktabs = TRUE) %>% 
  add_header_above(c("Variables(ref)" = 1,"Unadjusted" = 4, "Adjusted" = 4)) %>%
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
  
```

## Imputed model CCASAnet (Post ART)
```{r impu regression adjusted ccasanet post, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab13) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab13, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
    
``` 

## Imputed model VCCC (Post ART)
```{r impu 2 regression adjusted tn post, warning=FALSE, echo=FALSE, message=FALSE}

colnames <- data.frame( "RR","Lower .95 CI", "Upper .95 CI","P Value",stringsAsFactors = FALSE)
colnames(tab14) <- colnames


options(digits = 3)
options(scipen=999)
kbl(tab14, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
    
``` 


# KM Curves

## CCASAnet

```{r km CCASAnet, warning=FALSE, echo=FALSE, message=FALSE}
km_la <- la %>% filter(!is.na(histo_d)) %>% dplyr::select(patient_id,site,histo_d,death_d,last_d,art_sd)
km_la$death_y <- with(km_la,ifelse(!is.na(death_d),1,0))
km_la$time <- with(km_la,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(last_d),as.Date(histo_d),units ="days")/365.25))

tn$enrol_d <- ifelse(!is.na(tn$year_of_enrollment),
                     paste0(tn$year_of_enrollment, "-07-01"),NA)
tn$enrol_d <- as.Date(tn$enrol_d)
tn$death_d <- with(tn,ifelse(!is.na(age_at_death), enrol_d + (age_at_death - age_at_first_visit)*365.25,NA))
tn$age_last <- ifelse(tn$CENSOR_TYPE == "LAST VISIT", tn$age_at_last_visit,tn$age_at_censor)
tn$diff_enrol_last <- with(tn,(age_last - age_at_first_visit)*365.25)
tn$last_d <- with(tn,enrol_d + diff_enrol_last)
tn$last_d <- with(tn,ifelse(CENSOR_TYPE == "12/31/2021", as.Date("2021-12-31"),as.Date(last_d,origin="1970-01-01")))
tn$last_d <- as.Date(tn$last_d,origin="1970-01-01")
tn$last_d <- with(tn,ifelse(last_d > "2021-12-31",as.Date("2021-12-31"),last_d))
tn$last_d <- as.Date(tn$last_d,origin="1970-01-01")
tn$art_sd <- with(tn,enrol_d + (AGE_AT_ART_START - age_at_first_visit)*365.25)
tn$histo_d <- with(tn,enrol_d + (age_at_histo - age_at_first_visit)*365.25)
km_tn <- tn%>% filter(!is.na(histo_d)) %>% dplyr::select(CFAR_PID,histo_d,death_d,last_d,art_sd)
km_tn$death_y <- with(km_tn,ifelse(!is.na(death_d),1,0))
km_tn$death_d <- as.Date(km_tn$death_d,origin="1970-01-01")
km_tn$histo_d <- as.Date(km_tn$histo_d,origin="1970-01-01")
km_tn$time <- with(km_tn,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(last_d),as.Date(histo_d),units ="days")/365.25))


km1 <- survfit(Surv(time, death_y) ~ site, data=km_la)
km2 <- survfit(Surv(time, death_y) ~ 1, data=km_tn)

ggsurvplot(km1,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "CCASAnet",
                              ylab="Probability of death ",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))

la_pre <- km_la %>% filter(histo_d <= art_sd |  is.na(art_sd))
la_pre$pre_time <-  with(la_pre,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(art_sd),as.Date(histo_d),units ="days")/365.25))
la_post <- km_la %>% filter(histo_d> art_sd)
la_post$post_time <-  with(la_post,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(histo_d),as.Date(art_sd),units ="days")/365.25))


km1_pre <- survfit(Surv(pre_time, death_y) ~ site, data=la_pre)


ggsurvplot(km1_pre,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "CCASAnet pre-ART initiation",
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


km1_post <- survfit(Surv(post_time, death_y) ~ site, data=la_post)

ggsurvplot(km1_post,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              
                              title = "CCASAnet post-ART initiation",   
                              
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


```


## VCCC

```{r km VCCC, warning=FALSE, echo=FALSE, message=FALSE}

tn_pre <- km_tn %>% filter(histo_d <= art_sd | is.na(art_sd))
tn_pre$pre_time <-  with(tn_pre,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(art_sd),as.Date(histo_d),units ="days")/365.25))
tn_post <- km_tn %>% filter(histo_d> art_sd)
tn_post$post_time <-  with(tn_post,ifelse(death_y ==1,difftime(as.Date(death_d),as.Date(histo_d),units ="days")/365.25,
                                difftime(as.Date(histo_d),as.Date(art_sd),units ="days")/365.25))

ggsurvplot(km2,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "VCCC",
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))

km2_pre <- survfit(Surv(pre_time, death_y) ~ 1, data=tn_pre)


ggsurvplot(km2_pre,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "VCCC pre-ART initiation",
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


km2_post <- survfit(Surv(post_time, death_y) ~ 1, data=tn_post)

ggsurvplot(km2_post,pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "VCCC post-ART initiation",
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


```

## All sites

```{r km overall, warning=FALSE, echo=FALSE, message=FALSE}

fits <- list(tn= km2, la=km1)
km_la$site <- relevel(as.factor(km_la$site), ref = "brazil")
cox_model <- coxph(Surv(time, death_y) ~ site, data = km_la)
fit <- survfit(cox_model, newdata = data.frame(site = c("brazil", "chile","honduras","mexico","peru")))
surv_summary  <- summary(fit, times = 5)
five_year_survival <- as.data.frame(surv_summary$surv * 100)

cox_model2 <- coxph(Surv(time, death_y) ~ 1, data = km_tn)
fit1 <- survfit(cox_model2)
surv_summary  <- summary(fit1, times = 5)
five_year_survival2 <- as.data.frame(surv_summary$surv * 100)
five_year_survival <- merge(five_year_survival,five_year_survival2)
colnames(five_year_survival) <- c("Brazil","Chile","Honduras","Mexico","Peru","TN")
five_year_survival

ggsurvplot_combine(fits,km_la, pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,
                              legend="bottom",
                              legend.title="",
                              title = "All sites",
                              censor=FALSE,
                              palette = c("#000","#1A1A1A","gray80","#999","#666","#333333"),
                              linetype = c(1,4,1,6,5,9),
                              size = c(0.7,0.8,3,0.8,0.9,0.6),
                              legend.labs = c("Tennesse", "Brazil","Chile","Honduras","Mexico", "Peru"),
                              ylab="Probability of death ",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))

ggsurvplot_combine(fits,km_la, pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "All sites",
                              ylab="Probability of death ",
                              legend.labs = c("Tennesse", "Brazil","Chile","Honduras","Mexico", "Peru"),
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))

fits_pre <- list(tn= km2_pre, la=km1_pre)
ggsurvplot_combine(fits_pre,tn_pre, pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "All sites pre-ART initiation",
                              ylab="Probability of death ",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


fits_post <- list(tn= km2_post, la=km1_post)
ggsurvplot_combine(fits_post,tn_post, pval=TRUE,risk.table = FALSE,
                              conf.int=FALSE,censor=TRUE,
                              legend="bottom",
                              legend.title="",
                              title = "All sites post-ART initiation",
                              ylab="Probability of death diagnosis",
                              xlab="Years since Histoplasmosis diagnosis",
                              xlim= c(0, 20))


print(km1)
print(km1_pre)
print(km1_post)
print(km2)
print(km2_pre)
print(km2_post)


```




# CPH model summary

Notes 

+  HR for death 
+  ART status at Histo : if ART_sd < Histo_ d then it's POST-ART at histo, and if art_sd>= histo_d or missing then it’s PRE-ART at histo
+  Not using splines

### Table stratified by Site
```{r HR site}
label(c$age_at_diag) <- "Age at diagnosis"
label(c$art_status) <- "ART status wrt histo date"
ds <- c %>% dplyr::select(sex,age_at_diag,mig_ende,histo_y,cd4_diag_sq ,rna_diag_log ,site,death,art_status)
ds %>% 
  tbl_summary( 
     by = site, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2", "stat_3", "stat_4","stat_5","stat_6") ~ " Study Sites") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

### Table stratified by death for CCASAnet
```{r}
label(la_c$age_at_diag) <- "Age at diagnosis"
label(la_c$art_status) <- "ART status wrt histo date"
ds <- la_c %>% dplyr::select(sex,age_at_diag,mig_ende,histo_y,cd4_diag_sq ,rna_diag_log ,site,death,art_status)
ds %>% 
  tbl_summary( 
     by = death, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

### Table stratified by death for TN
```{r}
label(tn_c$art_status) <- "ART status wrt histo date"
label(tn_c$age_at_diag) <- "Age at diagnosis"
ds <- tn_c %>% dplyr::select(sex,age_at_diag,mig_ende,histo_y,cd4_diag_sq ,rna_diag_log ,site,death,art_status)
ds %>% 
  tbl_summary( 
     by = death, 
    statistic = list(all_continuous() ~ "{p25}  {median}  {p75}",
              all_categorical() ~ "{n} ({p}%)"),
              percent = c("column"),type = all_dichotomous() ~ "categorical"
              ) %>%
  add_n() %>%
  modify_header(label ~ "**Variable**") %>%
  modify_footnote(
    all_stat_cols() ~ "25th percentile Median 75th percentile or Frequency (%)"
  ) %>%
  bold_labels()

```

## Univariate analysis : CCASAnet 

```{r cox la, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame("HR","95% CI", "P Value",stringsAsFactors = FALSE)
colnames(cph_uni_tab_la) <- colnames


options(digits = 3)
options(scipen=999)
kbl(cph_uni_tab_la, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```

## Univariate analysis : TN

```{r cox tn, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "HR","95% CI", "P Value",stringsAsFactors = FALSE)
colnames(cph_uni_tab_tn) <- colnames


options(digits = 3)
options(scipen=999)
kbl(cph_uni_tab_tn, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```


## Multivariable analysis with stratification (including TN as a site)

```{r cox multi, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "Variable","HR","95% CI", "P Value",stringsAsFactors = FALSE)
colnames(cph_summ) <- colnames


options(digits = 3)
options(scipen=999)
kbl(cph_summ, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 
```


## Multivariable analysis with imputation (including TN as a site)

```{r cox multi imp, warning=FALSE, echo=FALSE, message=FALSE}
colnames <- data.frame( "Variable","HR","95% CI", "P Value",stringsAsFactors = FALSE)
colnames(cph_i) <- colnames
load(file="cph_i.Rdata")

options(digits = 3)
options(scipen=999)
kbl(cph_i, booktabs = TRUE) %>% 
 kable_styling(bootstrap_options = c("striped", "hover", "condensed")) 

cph_i <- cph_i %>%
  mutate(limits = str_extract(CI, "\\(([^)]+)\\)"))

cph_i <- cph_i %>%
  mutate(
    lower_limit = as.numeric(str_trim(sub("\\(([^,]+),.*", "\\1", limits))),
    upper_limit = as.numeric(str_trim(sub(".*,\\s*([^,]+)\\)", "\\1", limits)))
  )

cph_i$Effect1 <- log(cph_i$Effect)
cph_i$lower_limit1 <- log(cph_i$lower_limit)
cph_i$upper_limit1 <- log(cph_i$upper_limit)

cph_i$Effect1 <- round(cph_i$Effect1,digits=2)
cph_i$lower_limit1 <- round(cph_i$lower_limit1,digits=2)
cph_i$upper_limit1 <- round(cph_i$upper_limit1,digits=2)

cph_i <- cph_i[-c(1,3,9,24,25),]
cph_i$var <- c("Sex: Male",
                      "Migration status: Yes",
                      "Age at diagnosis (ref: 35)", "25",  "45", "55",
                      "Histo diagnosis year (ref: 2010)", "2000", "2005", "2015","2020",
                      "CD4 at histo diagnosis (ref: 50)", "10" ," 25", "100", "200",
                      "log HIV RNA at diagnosis (ref: 5)", "1","3","7", "ART Status at diagnosis: post initiation")


cph_i$var <- factor(cph_i$var, ordered = TRUE,levels = c("Sex: Male",
                      "Migration status: Yes",
                      "Age at diagnosis (ref: 35)", "25",  "45", "55", "65",
                      "Histo diagnosis year (ref: 2010)", "2000", "2005", "2015","2020",
                      "CD4 at histo diagnosis (ref: 50)", "10" ," 25", "100", "200",
                      "log HIV RNA at diagnosis (ref: 5)", "1","3","7","6", "ART Status at diagnosis: post initiation" ))

cph_i <- cph_i[c(3,4,5,6,1,7:21,2),]

cph_i <- cph_i %>%
  mutate(label = ifelse(is.na(Effect), "", sprintf("%.2f [%.2f, %.2f]", Effect, lower_limit, upper_limit)))


mylabs <- c(0.0156,0.0625,0.25,1,4)
library(gridExtra)

p1 <- ggplot(cph_i, aes(x=factor (var, level=var), y=cph_i$Effect1, ymin=cph_i$lower_limit1, ymax=cph_i$upper_limit1)) + 
  geom_linerange() +
  geom_hline(aes(x=0,yintercept=log(1)), lty=2) +
  geom_point(shape=18) +
   
  scale_y_continuous(name= " Hazard ratio 95% CI (log scale)",breaks = c(c(-6,-4,-2,0,2)),labels=c(mylabs)) +
  coord_flip() +
  theme(
    text = element_text(family = "Georgia"),
    plot.title = element_text(size = 14, color = "black", face = "bold"),
    plot.subtitle = element_text(face = "italic", color = "black", size = 14),
    plot.caption = element_text(face = "italic", size = 14, color = "black"),
    axis.title.x = element_text(face = "bold", size = 14, color = "black"),
    axis.text.x = element_text( size = 12, color = "black"),
    axis.text.y = element_text(size = 12, color = "black"),
    axis.title.y = element_blank(),
    panel.grid.major = element_line(color = "gray90"),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.background = element_rect(fill = "#fcfbfd"),
    plot.margin = unit(c(1, 1, 1, 0), "cm")) +
    scale_x_discrete(limits = rev, breaks = cph_i$var,expand = expansion(add = 0)) 
    
# Text column

p2 <- ggplot(cph_i, aes(y=factor (var, level=var), x = 1, label = label)) +
  geom_text(hjust = 0,size=3.6) + 
  scale_y_discrete(limits = rev, breaks = cph_i$var) +
  scale_x_continuous(expand = c(0, 0)) +
  theme_void() +
  theme(plot.margin = unit(c(0.6, 1, 2.3, -4), "cm"))
# Combine plots
forest <- grid.arrange(p1, p2, ncol = 2, widths = c(4, 1))
```