---
title: "Logistic Graph"
output:
  pdf_document: default
  word_document: default
editor: source
---

```{r}
#| include: false

rm(list=ls())
knitr::opts_chunk$set(echo = F,warning = F,message = F,
                      results = 'asis',
                      fig.width = 8,
                      fig.height = 8)
library(survminer)
library(arsenal)
library(Hmisc)
library(rms)
library(splines)
library(ggfortify)
library(patchwork)
library(gridExtra)
library(mice)
library(lubridate)
library(scales)
library(viridisLite)
library(tidyverse)
library(ggthemes)
library(foreach)

theme_set(
  theme_pander() +
    theme(
    legend.position = "bottom",
    title = element_text(size = 18),
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    axis.text = element_text(size = 14),
    legend.text = element_text(size = 14)
  )
)

scale_colour_discrete <-scale_colour_ptol
scale_color_discrete <-scale_colour_ptol

sur_trans = function() trans_new("sur",function(x) 1-x, function(x) 1-x)

tidydim <- function(x,prefix = "",surfix = "") {
  print(paste0(prefix,dim(x)[1],surfix))
  invisible(x)
}

source("fit.mult.impute.R")
```


## Enrollment as start
```{r}
set.seed(123123)
dt = readRDS("data/rebuilt_data.rds") %>% 
  ########### New analysis Only do HAART patieint#########
  filter(HAART>0) %>%
  select_if(function(x) length(na.omit(unique(x)))>1)
            
  #   ) %>% 
  # mutate(HAART = 1)
  ########### New analysis Only do HAART patieint#########

dt = select(dt,!where(lubridate::is.Date))%>% 
  mutate(day0 = futime<1,
         day1 = futime<=1,
         day14 = futime<=14,
         day30 = futime<=30,
         day180 = futime<=180,
         )

dd <- datadist(dt) 
options(datadist= 'dd')

dd$limits$baseage[2] <- 35
# dd$limits$male[2] <- "Male"
dd$limits$cd4_base[2] <- 350
dd$limits$baseyear[2] <- 2006

# set.seed(1)
# 
# dt_imp =
#   futuremice(dt,m = 20,n.core = 5,method = 'cart',maxit = 5)
# write_rds(dt_imp,"data/haart.rds")

dt_imp = read_rds("data/haart.rds")
```




```{r}

sites = unique(dt$site)

dt %>% 
  filter(tb==1) %>% 
  group_by(site,baseyear) %>% 
  count()

## Comment on mexico after 2019
```



```{r Peru}
model_Peru =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots =c(2012,2016,2019)) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2012,2016,2019)) +
       ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2012,2016,2019)) +
        ns(I(tb) * baseyear, knots =c(2014,2019)),
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2012,2016,2019)) +
        ns(I(tb) * baseyear, knots = c(2012,2016,2019))
      ,
      dt ,
      subset = (dt$site == 'Peru'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2012,2016,2019))  +
        ns(I(tb) * baseyear, knots =c(2014,2019))
      ,
    dt ,
    subset = (dt$site=='Peru'),
    family='binomial')
  )
```

```{r Honduras}
model_Honduras =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear,3) +
        ns(I(tb) * baseyear, 1),
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,3) 
      ,
    dt ,
    subset = (dt$site=='Honduras'),
    family='binomial')
  )
```



```{r Mexico}
model_Mexico =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2014))
      +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2014))
      +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear,knots = c(2011, 2014)) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day30 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2014)) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day180 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2014))+
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    )
  )
```



```{r Haiti}
model_Haiti =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2013,2018)) +
        ns(I(tb) * baseyear, knots = c(2013,2018))
      ,
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2015, 2019)) +
        ns(I(tb) * baseyear, knots = c(2011, 2015, 2019))
      ,
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2015, 2019)) +
        ns(I(tb) * baseyear, knots =c(2011, 2015, 2019)),
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2011,2015,2019)) +
        ns(I(tb) * baseyear, 1)
      ,
    dt ,
    subset = (dt$site=='Haiti'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2011,2015,2019))+
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Haiti'),
    family='binomial')
  )
```


```{r Brazil}
model_Brazil =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2010, 2014, 2019)) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2010, 2014, 2019)) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear, knots = c(2010, 2014, 2019)) +
        ns(I(tb) * baseyear, knots = c(2012, 2019)),
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2010,2014,2019)) +
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Brazil'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,knots = c(2011,2015,2019)) +
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Brazil'),
    family='binomial')
  )
```


```{r Overall}
model_Overall =
  list(
    glm(
      day0 ~ tb +
        ns(I(1 - tb) * baseyear, knots = c(2013,2018)) +
        ns(I(tb) * baseyear, knots = c(2013,2018))
      ,
      dt ,
      family = 'binomial'
    ),
    glm(
      day1 ~ tb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2015, 2019)) +
        ns(I(tb) * baseyear, knots = c(2011, 2015, 2019))
      ,
      dt ,
      family = 'binomial'
    ),
    glm(
      day14 ~ tb +
        ns(I(1 - tb) * baseyear, knots = c(2011, 2015, 2019)) +
        ns(I(tb) * baseyear, knots =c(2011, 2015, 2019)),
      dt ,
      family = 'binomial'
    ),
      glm(day30~ tb + 
        ns(I(1-tb)*baseyear,knots = c(2011,2015,2019)) +
        ns(I(tb) * baseyear, knots = c(2011,2015,2019))
      ,
    dt ,
    family='binomial'),
      glm(day180~ tb + 
        ns(I(1-tb)*baseyear,knots = c(2011,2015,2019))+
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    family='binomial')
  )
```


```{r here is the predicting matrix,eval=F}

#TODO: Only predicting exist data

dt_p = dt %>% 
  group_by(site,baselinetb,tb) %>% 
  reframe(baseyear = min(baseyear):max(baseyear)) %>% 
  ungroup() %>% 
  filter(baseyear<=2020)

dt_predict =
  foreach(s = c('Brazil','Honduras','Mexico','Peru','Haiti'),
          .combine = rbind) %do% {
    model = get(paste0("model_", s))
    result = 
      foreach(i =1:5, .combine = rbind) %do% {
        m = model[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
      }
    result
          }

dt_predict$day = fct_relevel(dt_predict$day,'day-0','day-1','day-7','day-30','day-180')
```

```{r eval=T}
dt_p = dt %>% 
  group_by(site,baselinetb,tb) %>% 
  reframe(baseyear = min(baseyear):max(baseyear)) %>% 
  ungroup()

dt_bra = foreach(i =1:5, .combine = rbind) %do% {
        s = 'Brazil'
        m = model_Brazil[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
      }
dt_hon = foreach(i =1:5, .combine = rbind) %do% {
        s = 'Honduras'
        m = model_Honduras[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
}
dt_mex = foreach(i =1:5, .combine = rbind) %do% {
        s = 'Mexico'
        m = model_Mexico[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
}
dt_per = foreach(i =1:5, .combine = rbind) %do% {
        s = 'Peru'
        m = model_Peru[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
}
dt_hai = foreach(i =1:5, .combine = rbind) %do% {
        s = 'Haiti'
        m = model_Haiti[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
}

dt_predict = do.call(rbind,
                     list(dt_bra,dt_hon,dt_mex,dt_per,dt_hai))%>% 
  filter(baseyear<=2020)

dt_predict$day = fct_relevel(dt_predict$day,'day-0','day-1','day-7','day-30','day-180')
```


```{r here is the predicting matrix for all}

#TODO: Only predicting exist data

dt_p = expand.grid(
  baseyear=unique(dt$baseyear),
  tb=(c(0,1)>0)
) %>% 
  filter(baseyear<=2020)

model = model_Overall
dt_all = 
      foreach(i =1:5, .combine = rbind) %do% {
        m = model[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p 
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
      }

dt_all$day = fct_relevel(dt_all$day,'day-0','day-1','day-7','day-30')
```


```{r}
#| fig-width: 8
#| fig-height: 10


dt_predict %>% 
  select(site,baseyear,baselinetb,day,y) %>% 
  arrange(site,baseyear,baselinetb) %>% 
  filter(baseyear %in% c(2006,2020),
         day %in% c('day-0','day-7'))

foreach(s = sites) %do%{
  ggplot(dt_predict %>% 
         filter(site==s),
       aes(x=baseyear,y=y))+
  geom_line(aes(color=baselinetb),size=1)+
  geom_ribbon(aes(ymin=lower,ymax=upper,group=as.factor(baselinetb)),alpha=0.2)+
  scale_x_continuous(breaks = seq(2006,2020,len=5))+
  facet_wrap(~day,nrow = 2) +
  labs(title = paste0("Freqency of ART Initiation:",s),
       y="Freqency",
       x="Year of Enrollment")+
  scale_color_discrete(name = "")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = -45))
}
```


```{r fig 3}
#| fig-width: 8
#| fig-height: 10

p = ggplot(dt_predict,
       aes(x=baseyear,y=y))+
  geom_line(aes(color=site),size=2,alpha=0.8)+
  scale_x_continuous(breaks = seq(2006,2020,len=4),
                     labels = round(seq(2006,2020,len=4)))+
  facet_wrap(day~baselinetb,nrow = 5) +
  labs(
    # title = "Freqency of ART Initiation",
       y="Estimated Probability of ART Initiation",
       x="Year of Enrollment")+
  scale_color_discrete(name = "")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = -45),
        axis.title.y = element_text(margin = margin(r = 15)))
plot(p)
ggsave("./figure/fig3_no_cen.png",p,dpi=400,width = 8,height = 10)
```


```{r}
dt_predict %>% 
  filter(day=='day-0',baseyear==2020,baselinetb=='No TB') %>% 
  select(site,baselinetb,y)

dt_predict %>% 
  filter(day=='day-0',baseyear==2018,baselinetb=='TB') %>% 
  select(site,baselinetb,y)

dt_predict %>% 
  filter(day=='day-7',baseyear==2020,baselinetb=='No TB') %>% 
  select(site,baselinetb,y)

dt_predict %>% 
  filter(day=='day-7',baseyear==2018,baselinetb=='TB') %>% 
  select(site,baselinetb,y)

dt_predict %>% 
  filter(day=='day-180',baseyear==2018,baselinetb=='No TB') %>% 
  select(site,baselinetb,y)

dt_predict %>% 
  filter(day=='day-180',baseyear==2018,baselinetb=='TB') %>% 
  select(site,baselinetb,y)
```


## CD4 <100

```{r cd4 less than 100}

dt %>% 
  group_by(baseyear) %>% 
  summarise(cd4_less_100 = mean(cd4_base<100,na.rm = T))

```


## HIV enrollment

```{r hiv dt}
set.seed(123123)
dt = readRDS("data/rebuilt_data.rds")

dt = select(dt,!where(lubridate::is.Date))%>% 
  filter(fuhiv2art>=0,
         hivyear>=2006,
         HAART>0
         ) %>% 
  mutate(futime = fuhiv2art,
         baseyear = hivyear,
         day0 = futime<1,
         day1 = futime<=1,
         day14 = futime<=14,
         day30 = futime<=30,
         day180 = futime<=180,
         ) 

dd <- datadist(dt) 
options(datadist= 'dd')

dd$limits$baseage[2] <- 35
# dd$limits$male[2] <- "Male"
dd$limits$cd4_base[2] <- 350
dd$limits$baseyear[2] <- 2006

```


```{r Peru 2}
model_Peru =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear,3) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear,3) +
       ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear,3) +
        ns(I(tb) * baseyear,1),
      dt ,
      subset = (dt$site == 'Peru'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Peru'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,3)  +
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Peru'),
    family='binomial')
  )
```

```{r Honduras 2}
model_Honduras =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear,3) +
        ns(I(tb) * baseyear, 1),
      dt ,
      subset = (dt$site == 'Honduras'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear,1)
      ,
      dt ,
      subset = (dt$site == 'Honduras'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,3) 
      ,
    dt ,
    subset = (dt$site=='Honduras'),
    family='binomial')
  )
```



```{r Mexico 2}
model_Mexico =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 1) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 1)
      +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear,2) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day30 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 2) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    ),
    glm(
      day180 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 2)+
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Mexico'),
      family = 'binomial'
    )
  )
```



```{r Haiti 2}
model_Haiti =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear,2) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear, 1),
      dt ,
      subset = (dt$site == 'Haiti'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear, 1)
      ,
    dt ,
    subset = (dt$site=='Haiti'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,3)+
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Haiti'),
    family='binomial')
  )
```


```{r Brazil 2}
model_Brazil =
  list(
    glm(
      day0 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
    glm(
      day1 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear, 1)
      ,
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
    glm(
      day14 ~ baselinetb +
        ns(I(1 - tb) * baseyear, 3) +
        ns(I(tb) * baseyear, 1),
      dt ,
      subset = (dt$site == 'Brazil'),
      family = 'binomial'
    ),
      glm(day30~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Brazil'),
    family='binomial'),
      glm(day180~ baselinetb + 
        ns(I(1-tb)*baseyear,3) +
        ns(I(tb) * baseyear,1)
      ,
    dt ,
    subset = (dt$site=='Brazil'),
    family='binomial')
  )
```


```{r here is the predicting matrix 2}

#TODO: Only predicting exist data

sites = unique(dt$site)

dt_p = dt %>% 
  group_by(site,baselinetb,tb) %>% 
  reframe(baseyear = min(baseyear):max(baseyear)) %>% 
  ungroup()%>% 
  filter(baseyear<=2020)

dt_predict =
  foreach(s = sites,
          .combine = rbind) %do% {
    model = get(paste0("model_", s))
    result = 
      foreach(i =1:5, .combine = rbind) %do% {
        m = model[[i]]
        d = c(0,1,7,30,180)[i]
      .dt_p = dt_p %>% filter(site==s)
      hat_y = predict(m, .dt_p, type = 'link')
      se_y = predict(m, .dt_p, type = 'link', se.fit = T)$se
      cbind(
        .dt_p,
        day = paste0("day-",d),
        y = 1 / (1 + 1 / exp(hat_y)),
        lower = 1 / (1 + 1 / exp(hat_y - 1.96 * se_y)),
        upper = 1 / (1 + 1 / exp(hat_y + 1.96 * se_y))
      )
      }
    result
          }

dt_predict$day = fct_relevel(dt_predict$day,'day-0','day-1','day-7','day-30')
```


```{r}
dt_predict %>% 
  filter(day=='day-0',baseyear==2020,baselinetb=='No TB') %>% 
  select(site,baselinetb,y) %>% 
  mutate(y = round(y,4))

dt_predict %>% 
  filter(day=='day-0',baseyear==2017,baselinetb=='TB') %>% 
  select(site,baselinetb,y) %>% 
  mutate(y = round(y,4))

dt_predict %>% 
  filter(day=='day-180',baseyear==2020,baselinetb=='No TB') %>% 
  select(site,baselinetb,y) %>% 
  mutate(y = round(y,4))

dt_predict %>% 
  filter(day=='day-180',baseyear==2017,baselinetb=='TB') %>% 
  select(site,baselinetb,y) %>% 
  mutate(y = round(y,4))
```

```{r}
p = ggplot(dt_predict,
       aes(x=baseyear,y=y))+
  geom_line(aes(color=site),size=2,alpha=0.8)+
  # geom_ribbon(aes(ymin=lower,ymax=upper,group=as.factor(site)),alpha=0.1)+
  scale_x_continuous(breaks = seq(2006,2020,len=4),
                     labels = round(seq(2006,2020,len=4)))+
  facet_wrap(day~baselinetb,nrow = 5) +
  labs(title = "Freqency of ART Initiation",
       y="Freqency",
       x="Year of Enrollment")+
  scale_color_discrete(name = "")+
  theme(plot.title = element_text(hjust = 0.5),
        axis.text.x = element_text(angle = -45))
plot(p)
```

